\bf Model Visualization Techniques for a Social Network Model

### 2017-04-11 paper.Rnw
par about geomnet

In \citeauthor{modelvis}, they define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. What does this definition mean for network data? For dynamic social networks, there are a few different data ``spaces." First, we have the actors and their corresponding covariates. Second, we have the dyad variables that describe the relationships between the variables. Finally, we have the time over which the network is allowed to evolve. These three data pieces can be visualized together in various ways. The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if the dyad between two nodes has value greater than zero, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side. These visual connections between the three data spaces are themselves part of the data space, because they can be viewed in different ways. Another way to connect the node and the edge data, as discussed in Section~\ref{sec:netvizintro}, is through the network's adjaceny matrix. The time dimension can also be incorporated with side-by-side adjacency matrix visualizations. 
\st{
\begin{itemize}
\item WHat is the model? the network model (SAOM model is just one of many)
\item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
\item Note: specifically dealing with dynamic networks observed at discrete time points. 
\end{itemize}
What are some things that could go in here? need more data examples; not just the friends data.
Harry Potter data??? I want to find my own data, too.}

>>>>

In \citeauthor{modelvis}, they define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. What does this definition mean for network data? For dynamic social networks, there are a few different data ``spaces." First, we have the actors and their corresponding covariates. Second, we have the dyad variables that describe the relationships between the variables. Finally, we have the time over which the network is allowed to evolve. These three data pieces can be visualized together in various ways. The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if the dyad between two nodes has value greater than zero, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side.

One tool that can bring these three data spaces together in this way is the \texttt{R} package \texttt{geomnet} \citep{geomnet}. Different aesthesic aspects of each piece of the node-link diagram can be tied to the underlying node or edge data. The color, size, and shape of the points can be used to represent variables in the node data, and the color, size, and linetype of the links between points can be used to represent the edge variables. In a social network, node data might be age, gender, and occupation of the person in the network, and node data might be length of connection between two people, how the people first met (school, work, church, etc.), and how often they interact. Pulling all of this information together with \texttt{geomnet} allows the entire data space to be viewed at once.

XXX INSERT AN EXAMPLE OF DATA WITH LOTS OF EDGE, NODE VARS OF INTEREST XXX

These visual connections between the node, edge, and time data spaces are themselves part of the data space, because they can be viewed in different ways. Another way to connect the node and the edge data, as discussed in Section~\ref{sec:netvizintro}, is through the network's adjaceny matrix. The time dimension can also be incorporated with side-by-side adjacency matrix visualizations. 
\st{
\begin{itemize}
\item WHat is the model? the network model (SAOM model is just one of many)
\item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
\item Note: specifically dealing with dynamic networks observed at discrete time points. 
\end{itemize}
What are some things that could go in here? need more data examples; not just the friends data.
Harry Potter data??? I want to find my own data, too.}
